PANCHANGE_data_preprocessing

### 2020-08-19 PANCHANGE_data_preprocessing2020-07-15.Rmd
synchronised gender vairable across cohorts

glad.dem.raw.id <- glad.dem.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         startDate.prepandemic = startDate, #Date when participant signed up to glad
         Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
         Gender = demographics.which_gender_do_you_identify_with, # Gender
         Sex = demographics.select_questionnaire_items_medical, #Sex
         EduYrs = demographics.school_education_school_years.txt, # Years in education
         demographics.college_or_university_degree, # level of education - univeristy
         demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
         demographics.o_levelsgcses_or_equivalent, # level of education - gcse
         demographics.cses_or_equivalent, # level of education - cse
         demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         Ethnicity = demographics.questions_based_ethnic_origin # Ethnicity
         ) %>%
  add_numeric(., exclude = exclude_cols_dem) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

  edgi.dem.raw.id <- edgi.dem.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
      startDate.prepandemic = startDate, #Date when participant signed up to edgi
      Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
      Gender = demographics.what_gender_do_you_identify_with, # Gender (which [GLAD] vs what [EDGI])
      Sex = demographics.select_questionnaire_items_birth, #Sex (different to GLAD)
      EduYrs = demographics.university_education_attend_school.txt, # Years in education (different to GLAD)
      demographics.college_or_university_degree, # level of education - univeristy
      demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
      demographics.o_levelsgcses_or_equivalent, # level of education - gcse
      demographics.cses_or_equivalent, # level of education - cse
      demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Ethnicity = demographics.what_is_your_ethnic_origin # Ethnicity (different to GLAD)
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

  nbr.dem.raw.id <- nbr.dem.raw %>%
    drop_na(subjectid) %>% # Drop NAs
    distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(subjectid, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
    separate(ID.intm, into = "ID", sep = 7) %>%
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
#      startDate.coping, # Date when participant signed up to nbr (is in the coping gad chunk)
#      Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
      Birthyear_unc = demographics.year, # Birthyear
      Birthmonth = demographics.month, #Birthmonth
      Birthday = demographics.day, #Birthday
      Gender = demographics.which_gender_do_you_identify_with, # Gender
      Sex = demographics.select_questionnaire_items_medical, #Sex
#      Ethnicity = demographics.questions_based_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

# Demographics variable creation
## Clean Age variable
Age outlier
```{r age outlier}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE ) {
# Define age limits 
age_upper_limit = 117 # The oldest person in the world is 117 years
age_lower_limit = 16

>>>>

glad.dem.raw.id <- glad.dem.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         startDate.prepandemic = startDate, #Date when participant signed up to glad
         Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
         Gender_unc = demographics.which_gender_do_you_identify_with, # Gender
         Sex = demographics.select_questionnaire_items_medical, #Sex
         EduYrs = demographics.school_education_school_years.txt, # Years in education
         demographics.college_or_university_degree, # level of education - univeristy
         demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
         demographics.o_levelsgcses_or_equivalent, # level of education - gcse
         demographics.cses_or_equivalent, # level of education - cse
         demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
         Ethnicity = demographics.questions_based_ethnic_origin # Ethnicity
         ) %>%
  add_numeric(., exclude = exclude_cols_dem) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

  edgi.dem.raw.id <- edgi.dem.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
      startDate.prepandemic = startDate, #Date when participant signed up to edgi
      Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
      Gender_unc = demographics.what_gender_do_you_identify_with, # Gender (which [GLAD] vs what [EDGI])
      Sex = demographics.select_questionnaire_items_birth, #Sex (different to GLAD)
      EduYrs = demographics.university_education_attend_school.txt, # Years in education (different to GLAD)
      demographics.college_or_university_degree, # level of education - univeristy
      demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
      demographics.o_levelsgcses_or_equivalent, # level of education - gcse
      demographics.cses_or_equivalent, # level of education - cse
      demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Ethnicity = demographics.what_is_your_ethnic_origin # Ethnicity (different to GLAD)
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

  nbr.dem.raw.id <- nbr.dem.raw %>%
    drop_na(subjectid) %>% # Drop NAs
    distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(subjectid, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
    separate(ID.intm, into = "ID", sep = 7) %>%
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
#      startDate.coping, # Date when participant signed up to nbr (is in the coping gad chunk)
#      Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
      Birthyear_unc = demographics.year, # Birthyear
      Birthmonth = demographics.month, #Birthmonth
      Birthday = demographics.day, #Birthday
      Gender_unc = demographics.which_gender_do_you_identify_with, # Gender
      Sex = demographics.select_questionnaire_items_medical, #Sex
#      Ethnicity = demographics.questions_based_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

## Demographics variable creation

# Gender categories synchronised

```{r gender recoding for GLAD}
if(GLAD == TRUE) {

dem.raw.id <- dem.raw.id %>%
  mutate(
    Gender = 
      recode_factor(Gender_unc,
        "Male" = "Male",
        "Female" = "Female",
        "Non-binary" = "Non-binary",
        "Prefer to self-define (please tell us more)" = "Prefer to self-define"
      )
  )
}
```

```{r gender recoding for EDGI}
if(EDGI == TRUE) {

dem.raw.id <- dem.raw.id %>%
  mutate(
    Gender = 
      recode_factor(Gender_unc,
        "Male" = "Male",
        "Female" = "Female",
        "Non-binary" = "Non-binary",
        "Prefer to self-define (please tell us more):" = "Prefer to self-define"
      )
  )
}
```

```{r gender recoding for NBR}
if(NBR == TRUE ) {

dem.raw.id <- dem.raw.id %>%
  mutate(
    Gender = 
      recode_factor(Gender_unc,
        "Male" = "Male",
        "Female" = "Female",
        "Non-binary" = "Non-binary",
        "Prefer to self-define (please tell us more)" = "Prefer to self-define"
      )
  )
}
```

# Clean Age variable
Age outlier
```{r age outlier}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE ) {
# Define age limits 
age_upper_limit = 117 # The oldest person in the world is 117 years
age_lower_limit = 16
